---
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    dev: png
    keep_md: yes
  word_document:
    toc: TRUE
    toc_depth: 3
  pdf_document:
    toc: TRUE
    toc_depth: 3
    dev: png
    extra_dependencies: ["float"]
    keep_md: yes
urlcolor: blue
title: '`r paste0(habitat," Indicator Quantile Report")`'
subtitle: 'SEACAR Analysis'
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
   warning=FALSE,
   message=FALSE,
   echo=FALSE,
   dpi=200
)
```



```{r libraries, message=FALSE, echo=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(data.table)
library(dplyr)
library(lubridate)
library(ggplot2)
library(ggpubr)
library(scales)
library(EnvStats)
library(tidyr)
library(kableExtra)
library(tidyverse)
library(stringr)
library(glue)
```

\newpage
# Overview

## Purpose


The purpose of the indicator quantiles is to flag records that are "unusual" relative to all of the data in the DDI for a given indicator in order to facilitate QA/QC. 
They are not used to filter any of the data for SEACAR analyses, and the presence of a LowerQuantile or UpperQuantile flag on a DDI record 
alone does not necessarily indicate there is any issue with the record (neither does the absence of a LowerQuantile or UpperQuantile flag 
necessarily mean that a data record is correct).


## Relevant file locations

Current values can be found in the "LowQuantile" and "HighQuantile" columns of the "Ref_Parameters" worksheet.

The R script described below and the output file can be found in the 
*FloridaSEACAR* *IndicatorQuantiles* repository on GitHub:  

* [https://github.com/FloridaSEACAR/IndicatorQuantiles](https://github.com/FloridaSEACAR/IndicatorQuantiles)


## Process steps  


**IQ_Report_Render.R & IQ_Report.Rmd**  

1.  The *IQ_Report_Render.R* script lists all files in a given directory and filters it to a list of DDI exports to 
evaluate considering a list of parameters to skip (user-defined).  
2.  User sets the desired upper and lower quantile thresholds, as well as a number of standard deviations away from the mean to use for the calculations.  
3.  User sets the string value(s) in the DDI exports that should be considered as NA values.  
4.  The remainder of the script loops through the file list, returning the values listed below and binding them together by row into a 
single Excel spreadsheet that is saved to the User's working directory.  
5.  For each habitat included in the User's working directory a PDF report will be created in the "output" folder using *IQ_Report.Rmd*,
which provides an overview of questionable / flagged values.  
6.  In addition to the PDF reports, each habitat will provide a .txt data output file in the "output/data" folder containing questionable values.


\newpage
# Summary

The following quantile thresholds are used for flagging "questionable" values:

* Lower quantile: **`r quant_low`**
* Upper quantile: **`r quant_high`**


Potential Included Parameters:
```{r, results='asis'}
cat("  \n\n")
for(param in sort(params_list)){
  if(param %in% parstoskip) next
  cat(paste0("* ", param))
  cat("\n")
}
```



```{r file_name display, results='asis'}
## display all file names for WC, or just one filename for others
cat("  \n\n")
if(habitat=="Water Column (Discrete)"){
  cat("The data files used for the analysis:  \n\n")
  for(file in short_files){
    cat(paste0("* **", file,"**"))
    cat("\n")
  }
} else {
  cat(paste0("The data file used for the analysis: **",file_short,"**"))
  cat("\n")
}

```


```{r table overview, results='asis', eval=TRUE}
options(knitr.kable.NA = "")
if(habitat=="Water Column (Discrete)"){
  qs_dat2 <- qs_dat %>%
    mutate(parameter = if_else(
      parameter == "Total Nitrogen",
      paste0("Nitrogen ", "(", sub_parameter, ")"),
      parameter
    ))
  table_data <- qs_dat2 %>%
    filter({{habitat}} == habitat) %>%
    select(parameter, sub_parameter, q_low, q_high, mean, n_tot, n_q_low, n_q_high) %>%
    mutate(`%_flagged` = (n_q_low + n_q_high) / n_tot * 100) %>%
    rename("sub_param" = "sub_parameter") %>%
    arrange(parameter)
} else if(habitat == "Submerged Aquatic Vegetation"){
  table_data <- qs_dat %>%
    filter({{habitat}} == habitat) %>%
    select(parameter, sub_parameter, q_low, q_high, mean, n_tot, n_q_low, n_q_high) %>%
    mutate(`%_flagged` = (n_q_low + n_q_high) / n_tot * 100) %>%
    rename("sub_param" = "sub_parameter") %>%
    arrange(parameter)
} else {
  table_data <- qs_dat %>%
    filter({{habitat}} == habitat) %>%
    select(parameter, q_low, q_high, mean, n_tot, n_q_low, n_q_high) %>%
    mutate(`%_flagged` = (n_q_low + n_q_high) / n_tot * 100) %>%
    arrange(parameter)
}

table_display <- kable(table_data,
      format="simple",
      caption="Indicator Quantile Overview",
      digits = 2) %>%
  kable_styling(latex_options="scale_down",
                position = "center")

if (any(str_detect(table_data$parameter, "(?i)Grazers and reef dependent species"))) {
  table_data$parameter <- str_replace(table_data$parameter,
                                      "(?i) - Grazers and reef dependent species$",
                                      " - *G&RDS*")
  
  table_display <- kable(table_data,
      format="simple",
      caption="Indicator Quantile Overview",
      digits = 2) %>%
  kable_styling(latex_options="scale_down",
                position = "center") %>%
    add_footnote("G&RDS: Grazers and reef dependent species  \n\n")
}

print(table_display)  

cat("\n\n")

```

*q_low*: Value corresponding to the qval_low quantile for the parameter in the DDI export.  

*q_high*: Value corresponding to the qval_high quantile for the parameter in the DDI export.  

*mean*: Mean value for the parameter in the DDI export.  

*n_tot*: Total number of records in the DDI export for the parameter.  

*n_q_low*: Number of records in the DDI export that are below q_low for the parameter.  

*n_q_high*: Number of records in the DDI export that are above q_high for the parameter.  

\%_*flagged*: Proportion of total records in the DDI export for the parameter which have been flagged as above q_high, or below q_low.  


```{r combined SAV 2}

flagged_combined_df <- flagged_combined_df %>%
  mutate(
    ParameterName = ifelse(
      habitat == "Submerged Aquatic Vegetation" & ParameterName %in% c("Percent Occurrence", "Percent Cover"),
      ifelse(
        CommonIdentifier %in% c("Total seagrass", "Total_SAV"),
        paste0(ParameterName, " - Total"),
        paste0(ParameterName, " - Species")
      ),
      ParameterName
    )
  )

```


```{r, quantile_table_display function, eval=TRUE}

quantile_table_display <- function(quantile) {
  q_values <- flagged_combined_df %>% filter(q_subset == quantile)

  params <- unique(q_values$ParameterName)
  
  for (p in params) {
    if(p %in% parstoskip) next

    if (nrow(q_values %>% filter(ParameterName == p)) > 0) {
      
      subtitle <- glue("## {p}")
      cat(subtitle, "\n\n")
      
      q_par_values <- q_values %>% 
        filter(ParameterName == p) %>%
        arrange(ProgramID, ProgramLocationID, SampleDate)
      
      if (habitat == "Coral Reef" | habitat == "Submerged Aquatic Vegetation" | habitat == "Water Column (Nekton)") {
        q_par_values <- q_par_values %>%
          select(RowID, ProgramID, ProgramName, ProgramLocationID, SampleDate, CommonIdentifier, ResultValue)
      } else {
        q_par_values <- q_par_values %>%
          select(RowID, ProgramID, ProgramName, ProgramLocationID, SampleDate, ResultValue)
      }
      
      flagged_program_id <- unique(q_par_values$ProgramID)
    
      q_value_table <- q_par_values %>%
        select(-ProgramName)
      
      if(habitat == "Submerged Aquatic Vegetation" & p %in% c("Percent Occurrence - Total", "Percent Cover - Total")){
        s <- str_split(p, " - Total")[[1]][1]
        low_thres <- qs_dat %>% filter(parameter==s, sub_parameter=="Total") %>% pull(q_low)
        high_thres <- qs_dat %>% filter(parameter==s, sub_parameter=="Total") %>% pull(q_high)
        
      } else if (habitat == "Submerged Aquatic Vegetation" & p %in% c("Percent Occurrence - Species", "Percent Cover - Species")){
        s <- str_split(p, " - Species")[[1]][1]
        low_thres <- qs_dat %>% filter(parameter==s, sub_parameter=="Species") %>% pull(q_low)
        high_thres <- qs_dat %>% filter(parameter==s, sub_parameter=="Species") %>% pull(q_high)
      } else {
        low_thres <- qs_dat %>% filter(parameter==p) %>% pull(q_low)
        high_thres <- qs_dat %>% filter(parameter==p) %>% pull(q_high)
      }
      
      threshold <- ifelse(quantile == "low", low_thres, high_thres)
      
      # Print data table
      q_data_table <- kable(q_value_table,
                            format="simple",
                            digits = 2,
                            caption=paste0("Flagged Values - ", str_to_title(quantile), " Indicator Quantile: **", round(threshold,5),"**")) %>%
        kable_styling(latex_options="scale_down",
                      position = "center")
      
      print(q_data_table)
      
      # List programs with flagged data below, with their ID
      cat("\n **Programs containing flagged data:** \n \n")
      
      for (p_id in flagged_program_id) {
        p_name <- unlist(unique(q_par_values %>% filter(ProgramID == p_id) %>% select(ProgramName)))
        cat(paste0("*",p_id,"*", " - ",p_name, "  \n"))
      }
      
      cat("  \n")
      cat("\\newpage")
      
    } else {
      cat(paste0("There are no *", str_to_title(quantile) ,"* Quantile Flagged Values for ", p))
    }
  
  }
}

```



```{r low quantiles, results='asis', eval=TRUE}

q_values <- flagged_combined_df %>% filter(q_subset == "low")
if (nrow(q_values) > 0) {
  cat("\\newpage")
  cat("# Low Quantile")
  cat("\n\n")
  quantile_table_display("low")
} else {
  cat(paste0("There are no Low Quantile flagged data entries for ", habitat))
}

```



```{r high quantiles, results='asis', eval=TRUE}

q_values <- flagged_combined_df %>% filter(q_subset == "high")
if (nrow(q_values) > 0) {
  cat("\\newpage")
  cat("# High Quantile")
  cat("\n\n")
  quantile_table_display("high")
} else {
  cat(paste0("There are no High Quantile flagged data entries for ", habitat))
}

```