---
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    dev: png
    keep_md: yes
  word_document:
    toc: TRUE
    toc_depth: 3
  pdf_document:
    toc: TRUE
    toc_depth: 3
    dev: png
    extra_dependencies: ["float"]
    keep_md: yes
urlcolor: blue
title: '`r paste0(h," Indicator Quantile Report")`'
subtitle: 'SEACAR Analysis'
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
   warning=FALSE,
   message=FALSE,
   echo=FALSE,
   dpi=200
)
```

\newpage
# Overview

## Purpose


The purpose of the indicator quantiles is to flag records that are "unusual" relative to all of the data in the DDI for a given indicator in order to facilitate QA/QC. 
They are not used to filter any of the data for SEACAR analyses, and the presence of a LowerQuantile or UpperQuantile flag on a DDI record 
alone does not necessarily indicate there is any issue with the record (neither does the absence of a LowerQuantile or UpperQuantile flag 
necessarily mean that a data record is correct).


## Relevant file locations

Current values can be found in the "LowQuantile" and "HighQuantile" columns of the "Ref_Parameters" worksheet.

The R script described below and the output file can be found in the 
*FloridaSEACAR* *IndicatorQuantiles* repository on GitHub:  

* [https://github.com/FloridaSEACAR/IndicatorQuantiles](https://github.com/FloridaSEACAR/IndicatorQuantiles)


## Process steps  


**IQ_Report_Render.R & IQ_Report.Rmd**  

1.  The *IQ_Report_Render.R* script lists all files in a given directory and filters it to a list of DDI exports to 
evaluate considering a list of parameters to skip (user-defined).  
2.  User sets the desired upper and lower quantile thresholds, as well as a number of standard deviations away from the mean to use for the calculations.  
3.  User sets the string value(s) in the DDI exports that should be considered as NA values.  
4.  The remainder of the script loops through the file list, returning the values listed below and binding them together by row into a 
single Excel spreadsheet that is saved to the User's working directory.  
5.  For each habitat included in the User's working directory a PDF report will be created in the "output" folder using *IQ_Report.Rmd*,
which provides an overview of questionable / flagged values.  
6.  In addition to the PDF reports, each habitat will provide a .txt data output file in the "output/data" folder containing questionable values.


\newpage
# Summary

The following quantile thresholds are used for flagging "questionable" values:

* Lower quantile: **`r quant_low`**
* Upper quantile: **`r quant_high`**  

The following parameters are being excluded from this analysis:

```{r excluded_params, results='asis'} 
for(par in parstoskip){
  cat(paste0("* *",par,"*"))
}
```

The following Programs are managed by RCP and will be highlighted throughout the report:

```{r highlight_rcp, results='asis'}
# Function to change text color for RCP programs
colorize <- function(x, color) {sprintf("\\textcolor{%s}{%s}", color, x)}

# Intended to highlight programs managed by RCP
# Skip=1 skips first row containing descriptor (SEACAR Program Matrix) 
rcp_ma <- read_excel("data/SEACAR Program Matrix.xlsx", skip=1)
rcp_id_list <- unique(rcp_ma$ID)
cat(colorize(rcp_id_list, "blue"), sep=", ")


```

**Included Indicators and Parameters and the files used in this analysis:**

```{r, results='asis'}

cat("  \n\n")
if(h=="Water Column"){
  
  for (type in names(data_directory[[h]])){
    cat(paste0("**",type,"**"))
    cat("  \n")
    indicators <- names(data_directory[[h]][[type]])
    
    for (ind in indicators){
      
      if(type=="Nekton"){
        ind_display <- paste0("*Indicator:* ",ind,"  -  *",file_short_list[[type]],"*")
      } else {
        ind_display <- paste0("*Indicator:* ",ind)
      }
      
      cat(ind_display)
      params <- names(data_directory[[h]][[type]][[ind]])
      
      cat("  \n")
      for (par in params){
        
        if(type=="Nekton"){
          par_display <- paste0("* ", par)
        } else if(type=="Discrete"){
          
          if(str_detect(par, "Total Nitrogen")){
            par_display <- paste0("* ", par, "\n     * *", file_short_list[[type]][[ind]][["Total Nitrogen"]], "*")
          } else {
            par_display <- paste0("* ", par, "\n     * *", file_short_list[[type]][[ind]][[par]], "*")
          }
          
        } else {
          # Continuous file display
          file <- str_replace(file_short_list[[type]][[ind]][[par]], "SW|SE|NW|NE", "*AllRegions*")
          
          par_display <- paste0("* ", par, "\n     * *", file[[1]], "*")
        }
        
        cat("\n")
        cat(par_display)
        cat("\n")
      }
      cat("  \n")
    }
  }
  
} else {
  
  cat("  \n")
  cat(paste0("*",file_short,"*"))
  cat(" \n\n")
  
  indicators <- names(data_directory[[h]])
  
  for(ind in indicators){
    params <- names(data_directory[[h]][[ind]])
    cat(paste0("*Indicator:* ",ind))
    cat("  \n")
    for(par in params){
      cat("\n")
      cat(paste0("* ", par))
      cat("\n")
    }
    cat("  \n")
  }
}
```



```{r file_name display, results='asis', eval=FALSE}
## display all file names for WC, or just one filename for others
cat("  \n\n")
if(h=="Water Column"){
  cat("The data files used for the analysis:  \n\n")
  
  for(type in names(file_short_list)){
    cat(paste0("**",type,"**"))
    cat("  \n")
    
    if(type=="Nekton"){
      cat("\n")
      cat(paste0("* ", file_short_list[[type]]))
      cat("  \n")
    } else {
      for(indicator in names(file_short_list[[type]])){
      
        cat(paste0("Indicator: ","*",indicator,"*"))
        cat("  \n")
        
        for(file in file_short_list[[type]][[indicator]]){
          cat("\n")
          
          if(type=="Continuous"){
            file <- str_replace(file, "SW|SE|NW|NE", "*AllRegions*")[[1]]
          }
          
          cat(paste0("* ", file))
          cat("\n")
        }
        cat("  \n")
      }
  }
  }
} else {
  cat(paste0("The data file used for the analysis: **",file_short,"**"))
  cat("\n")
}

```

\newpage
# Summary Tables

*q_low*: Value corresponding to the qval_low quantile for the parameter in the DDI export.  

*q_high*: Value corresponding to the qval_high quantile for the parameter in the DDI export.  

*mean*: Mean value for the parameter in the DDI export.  

*n_tot*: Total number of records in the DDI export for the parameter.  

*n_q_low*: Number of records in the DDI export that are below q_low for the parameter.  

*n_q_high*: Number of records in the DDI export that are above q_high for the parameter.  

\%_*flagged*: Proportion of total records in the DDI export for the parameter which have been flagged as above q_high, or below q_low.  


```{r table overview, results='asis', eval=TRUE}

options(knitr.kable.NA = "")

if(h=="Water Column"){
  for(type in names(data_directory[[h]])){
    type_title <- glue("## {type}")
    cat("\n\n")
    cat(type_title, "\n\n")
    
    indicators <- names(data_directory[[h]][[type]])
    
    for (ind in indicators){
      cat(paste0("*Indicator*: **",ind,"**"))
      cat("\n\n")
    
      table_data <- water_column_summary_directory[[type]] %>%
        filter(habitat == h, indicator == ind) %>%
        select(parameter, q_low, q_high, mean, n_tot, n_q_low, n_q_high) %>%
        mutate(`%_flagged` = (n_q_low + n_q_high) / n_tot * 100) %>%
        arrange(parameter)
    
      table_display <- kable(table_data,
                             format="simple",
                             caption="Indicator Quantile Overview",
                             digits = 2) %>%
        kable_styling(latex_options="scale_down",
                      position = "center")
    
      print(table_display)  
      
      cat("\n\n")
    }
    
  }
} else {
  
  # Latex format for longer oyster parameter names (quadsize included)
  table_format <- ifelse(h=="Oyster/Oyster Reef", "latex", "simple")
  select_columns <- c("parameter", "q_low", "q_high", "mean", "n_tot", "n_q_low", "n_q_high")
  
  for (ind in indicators){
    cat(paste0("*Indicator*: **",ind,"**"))
    cat("\n\n")
    
    table_data <- qs_dat %>%
      filter(habitat == h, indicator == ind) %>%
      select(select_columns) %>%
      mutate(`%_flagged` = (n_q_low + n_q_high) / n_tot * 100) %>%
      arrange(parameter)
    
    table_display <- kable(table_data,
                           format=table_format,
                           caption="Indicator Quantile Overview",
                           digits = 2,
                           booktabs = T) %>%
      kable_styling(latex_options=c("scale_down","HOLD_position"),
                    position = "center")
    
    print(table_display)  
    
    cat("\n\n")
  }
}

cat("\n")

```


```{r, quantile_table_display function, eval=TRUE}

quantile_table_display <- function(quantile, hab) {
  if (hab == "Water Column"){
    # List individual types (disc, cont, nekton) for Water Column habitat
    types <- names(data_directory[[hab]])
    
    for (type in types){
      if(type == "Nekton"){
        i <- "Nekton"
        cat("\n\n")
        
        # find parameters for each indicator
        parameters <- names(data_directory[[hab]][[type]][[i]])
        
        for (p in parameters) {
          
          # produce flagged data tables for given quantile (high or low)
          q_values <- data_directory[[hab]][[type]][[i]][[p]] %>% filter(q_subset == quantile)
          
          if(p %in% parstoskip) next
          
          if (nrow(q_values %>% filter(ParameterName == p)) > 0) {
            
            subtitle <- glue("### {p}")
            cat("\n\n")
            cat(subtitle, "\n\n")
            
            q_par_values <- q_values %>% 
              filter(ParameterName == p) %>%
              arrange(ProgramID, ProgramLocationID, SampleDate)
            
            # create mask for habitats that include speciesname(CommonID) info in table
            spec_habitats <- c("Coral/Coral Reef", "Submerged Aquatic Vegetation")
            
            if (h %in% spec_habitats) {
              q_par_values <- q_par_values %>%
                select(RowID, ProgramID, ProgramName, ProgramLocationID, SampleDate, CommonIdentifier, ResultValue)
            } else {
              q_par_values <- q_par_values %>%
                select(RowID, ProgramID, ProgramName, ProgramLocationID, SampleDate, ResultValue)
            }
            
            flagged_program_id <- unique(q_par_values$ProgramID)
            
            q_value_table <- q_par_values %>%
              mutate(rcp_cell_color = ifelse(ProgramID %in% rcp_id_list, "blue", "black")) %>%
              select(-ProgramName)
            
            q_value_table <- q_value_table %>%
              mutate(ProgramID = cell_spec(ProgramID, color=rcp_cell_color))
            
            low_thres <- water_column_summary_directory[[type]][parameter==p, q_low]
            high_thres <- water_column_summary_directory[[type]][parameter==p, q_high]
            
            threshold <- ifelse(quantile == "low", low_thres, high_thres)
            
            # Print data table
            q_data_table <- kable(q_value_table %>% select(-rcp_cell_color),
                                  format="simple",
                                  digits = 2,
                                  caption=paste0("Flagged Values - ", str_to_title(quantile), " Indicator Quantile: **",
                                                 round(threshold,5),"**")) %>%
              kable_styling(latex_options="scale_down",
                            position = "center")
            
            if(type=="Nekton"){
              print(q_data_table)
            }
            
            # List programs with flagged data below, with their ID
            cat("\n **Programs containing flagged data:** \n \n")
            
            for (p_id in flagged_program_id) {
              
              pid_display <- ifelse(p_id %in% rcp_id_list, colorize(p_id, "blue"), p_id)
              
              p_name <- unlist(unique(q_par_values %>% filter(ProgramID == p_id) %>% select(ProgramName)))
              cat(paste0("*",pid_display,"*", " - ",p_name, "  \n"))
            }
            
            cat("  \n")
            cat("\\newpage")
          } else {
            cat("  \n")
            cat(paste0("There are no *", str_to_title(quantile) ,"* Quantile Flagged Values for ", p))
            cat("  \n")
          }
        } 
      }
    }
    cat("  \n")
  } else {
    # List individual tables for each indicator
    indicators <- names(data_directory[[hab]])
    
    for (i in indicators){
      
      indicator_title <- glue("## Indicator: {i}")
      cat("\n\n")
      cat(indicator_title, "\n\n")
      
      # find parameters for each indicator
      parameters <- names(data_directory[[hab]][[i]])
      
      for (p in parameters) {
        
        # produce flagged data tables for given quantile (high or low)
        q_values <- data_directory[[hab]][[i]][[p]] %>% filter(q_subset == quantile)
        
        if(p %in% parstoskip) next
        
        if (nrow(q_values %>% filter(ParameterName == p)) > 0) {
          
          subtitle <- glue("### {p}")
          cat("\n\n")
          cat(subtitle, "\n\n")
          
          q_par_values <- q_values %>% 
            filter(ParameterName == p) %>%
            arrange(ProgramID, ProgramLocationID, SampleDate)
          
          # create mask for habitats that include speciesname(CommonID) info in table
          spec_habitats <- c("Coral/Coral Reef", "Submerged Aquatic Vegetation")
          
          if (h %in% spec_habitats) {
            q_par_values <- q_par_values %>%
              select(RowID, ProgramID, ProgramName, ProgramLocationID, SampleDate, CommonIdentifier, ResultValue)
          } else {
            q_par_values <- q_par_values %>%
              select(RowID, ProgramID, ProgramName, ProgramLocationID, SampleDate, ResultValue)
          }
          
          flagged_program_id <- unique(q_par_values$ProgramID)
          
          q_value_table <- q_par_values %>%
            select(-ProgramName)
          
          low_thres <- qs_dat %>% dplyr::filter(parameter==p) %>% dplyr::pull(q_low)
          high_thres <- qs_dat %>% dplyr::filter(parameter==p) %>% dplyr::pull(q_high)
          
          threshold <- ifelse(quantile == "low", low_thres, high_thres)
          
          # Print data table
          q_data_table <- kable(q_value_table,
                                format="simple",
                                digits = 2,
                                caption=paste0("Flagged Values - ", str_to_title(quantile), " Indicator Quantile: **",
                                               round(threshold,5),"**")) %>%
            kable_styling(latex_options="scale_down",
                          position = "center")
          
          print(q_data_table)
          
          # List programs with flagged data below, with their ID
          cat("\n **Programs containing flagged data:** \n \n")
          
          for (p_id in flagged_program_id) {
            
            pid_display <- ifelse(p_id %in% rcp_id_list, colorize(p_id, "blue"), p_id)
            
            p_name <- unlist(unique(q_par_values %>% filter(ProgramID == p_id) %>% select(ProgramName)))
            cat(paste0("*",pid_display,"*", " - ",p_name, "  \n"))
          }
          
          cat("  \n")
          cat("\\newpage")
        } else {
          cat("  \n")
          cat(paste0("There are no *", str_to_title(quantile) ,"* Quantile Flagged Values for ", p))
          cat("  \n")
        }
      }
    }
  }
}

```


```{r, WC_table display}

wc_program_table <- function(hab, type, i, p){
  # produce flagged data tables
  q_values <- data_directory[[hab]][[type]][[i]][[p]]
  
  data_summ <- water_column_summary_directory[[type]]
  
  if (nrow(q_values) > 0) {
    
    subtitle <- glue("### {p}")
    cat("\n\n")
    cat(subtitle, "\n\n")
    
    program_table <- q_values %>%
      dplyr::group_by(ProgramID, ProgramName) %>%
      dplyr::summarise(n_flagged = n(),
                       n_low = sum(q_subset=="low"),
                       n_high = sum(q_subset=="high"), .groups = "keep")
    
    # List of programs to show below table
    programs <- as.data.table(program_table[ , c("ProgramID", "ProgramName")])
    
    # Low and high quantile values
    low_thres <- data_summ[parameter==p, q_low]
    high_thres <- data_summ[parameter==p, q_high]
    
    # Total data count for display in program_table
    program_table$n_tot_all <- data_summ[parameter==p, n_tot]
    
    # Merge program data counts & create percentage (Program & All per parameter)
    program_data <- merge(program_table, program_counts[ParameterName==p & typeName==type, ], by = "ProgramID")
    program_data <- program_data %>%
      mutate(pct_prog = round(((n_flagged / n_tot)*100),3),
             pct_all = round(((n_flagged / n_tot_all)*100),3)) %>%
      select(ProgramID, n_flagged, n_low, n_high, n_tot, n_tot_all, pct_prog, pct_all)
    
    program_data <- program_data %>%
      dplyr::rowwise() %>%
      dplyr::mutate(cell_color = ifelse(pct_prog > 1, "orange", "black"),
                    rcp_cell_color = ifelse(ProgramID %in% rcp_id_list, "blue", "black")) %>%
      dplyr::mutate(pct_prog = cell_spec(pct_prog, color=cell_color),
                    ProgramID = cell_spec(ProgramID, color=rcp_cell_color)) %>%
      select(-c(cell_color, rcp_cell_color))
    
    cat(paste0(
      "Quantile Values:  \n",
      "*Low*: **", round(low_thres,5), "**\n",
      "*High*: **", round(high_thres,5), "**\n",
      "  \n",
      "Total data for *",p,"* across all programs: **", unique(program_data$n_tot_all),"**" 
    ))
    cat("  \n")
    
    # Print data table
    program_kable <- kable(program_data %>% select(-n_tot_all),
                           format="simple",
                           digits = 3,
                           booktabs = T,
                           caption=paste0("Flagged data, overview by program for **", p,"**")) %>%
      kable_styling(latex_options=c("scale_down", "HOLD_position"),
                    position = "center")
    
    print(program_kable)
    
    # List programs with flagged data below, with their ID
    cat("\n **Programs containing flagged data:** \n \n")
    
    for (p_id in unique(programs$ProgramID)) {
      pid_display <- ifelse(p_id %in% rcp_id_list, colorize(p_id, "blue"), p_id)
      p_name <- programs[ProgramID==p_id, ProgramName]
      cat(paste0("*",pid_display,"*", " - ",p_name, "  \n"))
    }
    # Add space between each parameter listing
    cat("  \n")
    cat("  \n")
    cat("  \n")
  } else {
    cat("  \n")
    cat(paste0("There are no *", str_to_title(quantile) ,"* Quantile Flagged Values for ", p))
    cat("  \n")
  }
}

```


\newpage

```{r WC Formatting, results='asis'}

if(h=="Water Column"){
  for(type in names(data_directory[[h]])){
    cat("\n\n")
    cat("\\newpage")
    cat(glue("# {type}"), "\n\n")
    
    # Discrete & Continuous display Program tables
    if(!type=="Nekton"){
      for(i in names(data_directory[[h]][[type]])){
        
        cat("\n\n")
        cat(glue("## {type} - {i}"), "\n\n")
        
        for(p in names(data_directory[[h]][[type]][[i]])){
          cat("\n\n")
          wc_program_table(h, type, i, p)
        }
      }
    } else {
      # Nekton
      cat("## Low Quantile")
      cat("\n\n")
      quantile_table_display("low", h)
      
      cat("\\newpage")
      cat("  \n")
      cat("## High Quantile")
      cat("\n\n")
      quantile_table_display("high", h)
    }
  }
}


```


```{r low quantiles, results='asis', eval=TRUE}

if(!h=="Water Column"){
  cat("\\newpage")
  cat("# Low Quantile")
  cat("\n\n")
  quantile_table_display("low", h)
}

```



```{r high quantiles, results='asis', eval=TRUE}

if(!h=="Water Column"){
  cat("\\newpage")
  cat("# High Quantile")
  cat("\n\n")
  quantile_table_display("high", h)
}

```



```{r data flag check, results='asis'}

check_flag <- function(h){
  data_combined_subset <- data.table()
  if(!h=="Water Column"){
    for(i in names(data_directory[[h]])){
      data_combined_subset <- rbind(data_combined_subset, plyr::rbind.fill(data_directory[[h]][[i]]), fill=TRUE)
    }
    
    # Compute statistics grouped by ParameterName
    flag_table <- data_combined_subset[,
                                       .(n_high = sum(q_subset == "high"),
                                         n_high_flagged = sum(q_subset=="high" & str_detect(SEACAR_QAQCFlagCode, "17Q")),
                                         n_low = sum(q_subset == "low"),
                                         n_low_flagged = sum(q_subset=="low" & str_detect(SEACAR_QAQCFlagCode, "16Q"))),
                                       by = ParameterName]
    
  } else {
    for(type in names(data_directory[[h]])){
      for(i in names(data_directory[[h]][[type]])){
        data2 <- plyr::rbind.fill(data_directory[[h]][[type]][[i]])
        data2$type_name <- type
        data2$indicator <- i
        
        data_combined_subset <- rbind(data_combined_subset, data2, fill=TRUE)
      }
    }
    
    # Compute statistics grouped by ParameterName
    flag_table <- data_combined_subset[, 
                                       .(n_high = sum(q_subset == "high"),
                                         n_high_flagged = sum(q_subset=="high" & str_detect(SEACAR_QAQCFlagCode, "17Q")),
                                         n_low = sum(q_subset == "low"),
                                         n_low_flagged = sum(q_subset=="low" & str_detect(SEACAR_QAQCFlagCode, "16Q")),
                                         indicator = unique(indicator)), 
                                       by = list(ParameterName, type_name)]
  }
  
  # add signifier to any entry that doesn't have correct flag
  # Color coding cells
  flag_table <- flag_table %>%
    dplyr::rowwise() %>%
    dplyr::mutate(cell_color_low = ifelse(n_low == n_low_flagged, "black", "red"),
                  cell_color_high = ifelse(n_high == n_high_flagged, "black", "red")) %>%
    dplyr::mutate(n_low = cell_spec(n_low, color = cell_color_low),
                  n_low_flagged = cell_spec(n_low_flagged, color = cell_color_low),
                  n_high = cell_spec(n_high, color = cell_color_high),
                  n_high_flagged = cell_spec(n_high_flagged, color = cell_color_high)) %>%
    select(-c(cell_color_high, cell_color_low))
  
  names(flag_table) <- gsub("_", " ", names(flag_table))
  
  # produce table where quantile flags may not be correctly applied
  different_codes <- data_combined_subset[
    (q_subset=="high" & !str_detect(SEACAR_QAQCFlagCode, "17Q")) | 
      (q_subset=="low" & !str_detect(SEACAR_QAQCFlagCode, "16Q")),
    c("RowID","ProgramID","ParameterName","ResultValue","SEACAR_QAQCFlagCode","SEACAR_QAQC_Description","q_subset")]
  
  flag_display_table <- kable(flag_table,
                              format="latex",
                              digits = 2,
                              escape = F,
                              booktabs=T) %>%
    kable_styling(latex_options=c("scale_down", "HOLD_position"),
                  position = "center")
  
  print(flag_display_table)
  
  if(nrow(different_codes)>0){
    cat("  \n")
    cat("Entries where ResultValue is above or below quantile, but expected SEACAR_QAQCFlagCode is not being applied")
    cat("  \n")
    
    code_display_table <- kable(different_codes %>% select(-SEACAR_QAQC_Description) %>% arrange(ProgramID),
                                format="latex",
                                digits = 4,
                                caption="SEACAR QAQC Flag Code discrepancies",
                                longtable = T,
                                booktabs = T) %>%
      kable_styling(latex_options=c("HOLD_position","repeat_header"),
                    position = "center",
                    font_size = 8)
    
    print(code_display_table)
    cat("  \n")
    for(desc in unique(different_codes$SEACAR_QAQC_Description)){
      cat(paste0("* ",desc))
      cat("  \n")
    }
  }
}

cat("# QAQC Quantile Flag Check")
cat("  \n")
cat(paste0("* n_high is the amount of data above the quantile value.   \n * n_high_flagged is the amount of data above the quantile value AND containing proper SEACARQAQCFlag of 17Q (ResultValue above quantile value)."))
cat("  \n")
cat(paste0("* n_low is the amount of data below the quantile value.   \n * n_low_flagged is the amount of data below the quantile value AND containing proper SEACARQAQCFlag of 16Q (ResultValue below quantile value)."))
cat("  \n")
cat("* If everything is in order, these values should be the same. Any discrepancies therein should be fastidiously noted.")
cat("  \n")

check_flag(h)


```


```{r expected data check function}

expected_data_check <- function(h){
  params <- expected_data[habitat==h, unique(ParameterName)]
  
  # dynamically include the correct text for each parameter
  included_text <- function(param){
    if(str_detect(param, "Braun")){
      text_string <- "Expected values for **Braun Blanquet Score**: *(0, 0.1, 0.5, 1, 2, 3, 4, 5)*"
    }
    if(str_detect(param, "Modified Braun")){
      text_string <- "Expected values for **Modified Braun Blanquet Score**: *(0, 0.1, 0.5, 1, 2, 3, 4, 5)*"
    }
    if(str_detect(param, "Percent")){
      text_string <- "Expected **Percent Cover** and **Percent Occurrence** values: *0-100 (inclusive)*"
    }
    if(str_detect(param, "Presence")){
      text_string <- "Expected **Presence/Absence** values: *0 or 1*"
    }
    if(str_detect(param, "Secchi")){
      text_string <- "Number of **Secchi Depth** readings with *SEACAR_QAQCFlagCode* of *8Q* (Original Value = 'ON BOTTOM'. Replaced with Total Depth)"
    }
    return(cat(paste0("  \n" , text_string, "  \n")))
  }
  
  for(p in params){
    included_text(p)
    # Secchi is checking for ValueQualifier, all else look at SEACAR_QAQCFlagCode
    flag_var <- "SEACAR_QAQCFlagCode"
    flag_val <- ifelse(str_detect(p, "Secchi"), "8Q", "15Q")
  }
  
  # Table of unexpected data sums, with flag sums
  flag_table <- expected_data[habitat==h, ] %>% 
    dplyr::group_by(ParameterName) %>% 
    dplyr::summarise(n = n(),
                     n_flagged = sum(str_detect((!!as.name(flag_var)), flag_val)))
  
  # Apply color if the 2 sums differ
  color_flag_table <- flag_table %>%
    dplyr::rowwise() %>%
    dplyr::mutate(cell_color = ifelse(n == n_flagged, "black", "red")) %>%
    dplyr::mutate(n = cell_spec(n, color=cell_color),
                  n_flagged = cell_spec(n_flagged, color=cell_color)) %>%
    select(-cell_color)
  
  names(color_flag_table) <- gsub("_", " ", names(color_flag_table))
  
  # Display table in report
  flag_display_table <- kable(color_flag_table,
                              format="latex",
                              digits=2,
                              escape=F,
                              booktabs=T) %>%
    kable_styling(latex_options="HOLD_position",position = "center")
  
  print(flag_display_table)
  
  cat("  \n")
  
  
  for(p in params){
    
    cat(glue("## {p}"))
    cat("  \n")
    
    results_overview <- expected_data[habitat==h, ] %>% 
      dplyr::filter(ParameterName==p) %>%
      dplyr::group_by(ProgramID, ProgramName) %>%
      dplyr::summarise("n data" = n())
    
    results_overview_kable <- kable(results_overview,format="latex",escape=F, booktabs=T) %>%
      kable_styling(latex_options="HOLD_position", position="center")
    
    print(results_overview_kable)
    cat("  \n")
    
  }

}

```



```{r expected data check, results='asis'}

if(nrow(expected_data)>0){
  if(nrow(expected_data[habitat==h, ])>0){
    
    cat("\\newpage")
    cat("# Expected Values Check")
    cat("  \n")
    cat("Attempts to uncover any unexpected ResultValues. Highlights any values that are not expected and compares 
    them with the intended SEACAR_QAQCFlagCode of **15Q** - *Unexpected Value: Outside expected parameter values*
    or ValueQualifier of **S** for Secchi Depth (visible on Bottom)  \n")
    cat("    \n")
    cat("  \n")
    cat("* *n* is the amount of entries containing unexpected values    \n")
    cat("  \n")
    cat("* *n flagged* is the amount of entries containing the intended SEACAR_QAQCFlagCode of **15Q**    \n")
    cat("  \n")
    cat("* The following Parameters are checked and their results will be listed below if there are any unexpected values:\n")
    cat("  + Presence/Absence    \n")
    cat("  + Secchi Depth Visible on Bottom    \n")
    cat("  + Braun Blanquet    \n")
    cat("  + Modified Braun Blanquet    \n")
    cat("  + Percent Cover    \n")
    cat("  + Percent Live Tissue    \n")
    cat("  + Total/Canopy Percent Cover    \n")
    cat("  \n")
    cat("  \n")
    
    expected_data_check(h)
    cat("  \n")
  }
} else {
  cat("\\newpage")
  cat("# Expected Values Check")
  cat("  \n")
  cat("Congratulations! There are *no* unexpected values in this dataset.    \n")
}


```


```{r SAV-BB Expected Data Check, eval=FALSE, results='asis'}

if(h=="Submerged Aquatic Vegetation"){
  cat("\\newpage")
  cat("# Expected Values Check")
  cat("  \n")
  cat("Attempts to uncover irregularities with regard to Braun Blanquet and Modified Braun Blanquet Scores. Highlights any values that are not 
      *(0, 0.1, 0.5, 1, 2, 3, 4, 5)* and compares them with the intended SEACAR_QAQCFlagCode of **15Q** - *Unexpected Value: Outside expected 
      parameter values*    \n")
  cat("    \n")
  cat("* *n* is the amount of entries containing unexpected values    \n")
  cat("  \n")
  cat("* *n flagged* is the amount of entries containing the intended SEACAR_QAQCFlagCode of **15Q**    \n")
  cat("  \n")
  
  # Table of unexpected data sums, with flag sums
  flag_table <- expected_data %>% 
    dplyr::group_by(ParameterName) %>% 
    dplyr::summarise(n = n(),
                     n_flagged = sum(str_detect(SEACAR_QAQCFlagCode, "15Q")))
  
  # Apply color if the 2 sums differ
  color_flag_table <- flag_table %>%
    dplyr::rowwise() %>%
    dplyr::mutate(cell_color = ifelse(n == n_flagged, "black", "red")) %>%
    dplyr::mutate(n = cell_spec(n, color=cell_color),
                  n_flagged = cell_spec(n_flagged, color=cell_color)) %>%
    select(-cell_color)
  
  names(color_flag_table) <- gsub("_", " ", names(color_flag_table))
  
  # Display table in report
  flag_display_table <- kable(color_flag_table,
                              format="latex",
                              digits=2,
                              escape=F,
                              booktabs=T) %>%
    kable_styling(latex_options="HOLD_position",position = "center")
  
  print(flag_display_table)
    
  for(p in unique(expected_data$ParameterName)){
    cat(glue("## {p}"))
    cat("  \n")
    
    results_overview <- expected_data %>% dplyr::filter(ParameterName==p) %>%
      dplyr::group_by(ProgramID, ProgramName) %>%
      dplyr::summarise("n data" = n())
    
    results_overview_kable <- kable(results_overview,format="latex",escape=F, booktabs=T) %>%
      kable_styling(latex_options="HOLD_position", position="center")
    
    print(results_overview_kable)
    cat("  \n")
    
  }
}

```